### 2021-03-31 README.Rmd
update readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "World Heritage Sites on Wikipedia: Cultural Heritage Activism in a Context of Constrained Agency"
```

# worldheritagewikipedia

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/worldheritagewikipedia/master?urlpath=rstudio) [![Knit-Rmd-in-Docker](https://github.com/benmarwick/worldheritagewikipedia/workflows/.github/workflows/knit-Rmd-in-docker.yaml/badge.svg)](https://github.com/benmarwick/worldheritagewikipedia/actions)

This repository contains the data and code for our paper:

> Marwick, B and P. Smith, (2021). `r Title`. _Big Data and Society_ <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Marwick, B and P. Smith, (2021). `r Title`.  Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.31235/osf.io/7gvza>

### How to cite

Please cite this compendium as:

> Marwick, B and P. Smith, (`r format(Sys.Date(), "%Y")`). Compendium of R code and data for `r Title`. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.17605/OSF.IO/AY27G>

## Contents

The **analysis** directory contains:

- [:file_folder: paper](/analysis/paper): R Markdown source document for manuscript and R script files. Includes R code to reproduce the figures and tables generated by the analysis. It also has a rendered version, `paper.docx`, suitable for reading (the code is replaced by figures and tables in this file)    
- [:file_folder: data](/analysis/data): Data used in the analysis.  
- [:file_folder: figures](/analysis/figures): Plots and other illustrations using in the paper

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/benmarwick/worldheritagewikipedia/archive/master.zip>

Or you can install this compendium as an R package, worldheritagewikipedia, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
remotes::install_github("benmarwick/worldheritagewikipedia")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.